Peer-to-Peer Permissionless Consensus via Reputation

### 2023-08-14 freechains/rebuttal.md

> Quanto a solução, senti falta de alguma discussão sobre as limitações.
> Em especial, a falta de finalidade, i.e., um post pode ser excluído devido
> aos forks (no momento do merge).
> Este post poderia ser de um usuário legitimo ? quanto isso ocorre e quais as
> implicações?

Sim, para isso, ele teria que escrever em um fork "obscuro", desconectado do
resto dos usuários legítimos.
Dessa forma, ao ser reconectado, esse fork poderia sim ter posts removidos.
A única implicação é que ele teria que postar novamente a mesma mensagem.

No meio da Seção 2.1, essa possibilidade é levantada, mas não com tanta ênfase:

-------------------------------------------------------------------------------

@@

A Figura 3 já tem essa transição.
Há uma seta bi-direcional entre os dois estados.
(Espero ter entendido corretamente esse comentário.)

# Reviewer 2

> It is strange that known techniques like EigenTrust are not even mentioned,
> even though we are in the blockchain domain.

We added a new paragraph at the end of Section 4 (Related Work), as follows:

> Some works for P2P file sharing propose to account the reputation of peers to
> form a web of trust based on their behavior
> history~\cite{p2p.rep.wang,p2p.rep.eigentrust}.
> However, there are three key aspects in terms of scope that distinguishes our
> reputation system: its focus on the contents, its subjective nature, and its
> dependency on consensus.
> First, it is the actual contents that are stored and evaluated in the forums,
> with the identities being a secondary attribute.
> Second, content evaluation is subjective, as well as any decision to revoke
> posts or to fork forums.
> Third, at any given point in time, peers must agree on a common forum DAG
> prefix with global and deterministic accuracy to act identically as a whole.
> In this case, consensus is fundamental, since a subtle off-by-one discrepancy
> in reputation creates an irreconcilable fork in the network.
> Achieving consensus on a permissionless network is the key contribution of this
> work.

-------------------------------------------------------------------------------

### 2023-08-14 freechains/freechains.tex
...

In this work, we adapt the idea of scarce resources to reach consensus and
eradicate Sybils, but in the context of social content sharing.
We indicate 4 main contributions:
%
\begin{enumerate}
\item Recognizing published contents as scarce resources, since they require
human work.
Work is manifested as new posts, which if approved by others, reward authors
with reputation tokens, which are further used to evaluate other posts with
likes and dislikes.
With such \emph{proof-of-authoring} mechanism, token generation is expensive,
while verification is cheap and shared by multiple users.
%
\item Using the reputation system to determine consensus.
Due to decentralization and network partitions, posts in a timeline form a
causal graph with only partial order, which we promote to a total order based
on the reputation of authors.
The consensus order is fundamental to detect conflicting operations, such as
likes with insufficient reputation (akin to Bitcoin's double spending).
%
\item Allowing users to create diversified forums of interest (not a
singleton blockchain), each as an independent timeline with its own
subjective authoring etiquette.
%
\item Supporting content removal without compromising the integrity of the
decentralized blockchain.
Users have the power to revoke posts with dislikes, and peers are forced to
remove payloads, only forwarding associated blockchain metadata.
\end{enumerate}
%
We integrated the proposed consensus algorithm into \FC~\cite{fcs.sbseg20}, a
practical peer-to-peer (P2P) content dissemination protocol that provides
strong eventual consistency~\cite{p2p.sec}.
%p2p.crdts,
To show the feasibility of the consensus mechanism, we simulated months of
activity of a chat channel and years of a newsgroup forum, both extracted from
publicly available Internet archives.

@@

In the absence of moderation, permissionless P2P forums are impractical, mostly
because of Sybils abusing the system.
For instance, it should take a few seconds to generate thousands of fake
identities and SPAM millions of messages.
%Hence, without moderation, there are no limits on the number and size of posts
%and no reasonable policy to distinguish quality.
For this reason, we propose a reputation system that works together with a
consensus algorithm to resist Sybil attacks.
% and make peer-to-peer public forums practical.
We consider Sybils to be groups of throw-away identities and machines with no
previous reputation in the forums.

@@

Note that even non-malicious network partitions may experience conflicts when
merging branches.
For instance, dislike operations in one branch to past posts in the common
prefix could possibly block posts from users in the other branch.
%, depending on how the posts are ordered.
Therefore, we need to order branches with a global consensus, such that we can
validate operations consistently across the network.
%
Note also that, to reach consensus, we should preferably rely on the maximum
information that all branches agree on, which is exactly the posts in the
common prefix.

@@

As a counterpoint, maybe users $a$ and $b$ have abandoned the forum for months,
and thus \code{branch-1} is actually legitimate.
In this case, users $a$ and $b$ might be the ones trying to take over the
forum and would succeed, since they control the majority of \reps in the common
prefix.
Yet another possibility is that both branches are legitimate but became
disconnected for a long period of time.
In any case, it is unacceptable that an old remote branch affects a local
active forum.

For this reason, the consensus algorithm includes an extra constraint that
prevents long-lasting local branches to merge, which leads to \emph{hard forks}
in the network.
A hard fork occurs when a local branch crosses a predetermined and irreversible
threshold of \emph{7 days} or \emph{100 posts} of activity.
In this case, regardless of the remote branch reputation, the local branch
takes priority and is ordered first.
This situation is analogous to a hard fork in Bitcoin and the branches become
incompatible and will never synchronize again.
More than simple numeric disputes, a hard fork represents a social conflict in
which reconciling branches is no longer possible.

@@

In summary, the rules to merge a branch \code{j} from a remote machine into a
branch \code{i} from a local machine are as follows:
\begin{itemize}
    \item \code{i} is first if it crosses the activity threshold of
          \emph{7 days} or \emph{100 posts}, regardless of \code{j}.
    \item \code{i} or \code{j} is first, whichever has more reputation in the
          common prefix, as discussed in Section~\ref{sec.design.basic}.
    \item otherwise, branches use the arbitrary criteria of lexicographical
          order of the post hashes immediately after the common prefix.
\end{itemize}

@@

In order to support content moderation and mitigate abuse, we integrated the
proposed reputation and consensus mechanism of Section~\ref{sec.design} with
\FC.
Table~\ref{fig.rules} details the concrete reputation rules we conceived for
\FC, which is compatible with the general rules of Table~\ref{fig.general}, and
which are discussed as follows through an example.
Authors have to sign posts in order to be accounted by the reputation system
and operate in the chains.
We start by creating an identity whose public key is assigned as the pioneer in
a chain named \emph{forum}:

@@

%already in the community vouch for new users, or (ii) imposing explicit costs
%for new posts, such as proof of work.
%The most basic concern in public forums is to resist Sybil attacks.
%Fully P2P systems cannot rely on logins or {\footnotesize CAPTCHAs} due to the
%lack of a central authority.
%Viable alternatives include (i) building social trust graphs, in which users
%already in the community vouch for new users, or (ii) imposing explicit costs
%for new posts, such as proof of work.
%
%We propose a mix between trust graphs and economic costs.
%
In order to resist Sybils, we propose a mix between building social trust
graphs and imposing explicit costs for new posts:
Rule~\code{4.a} imposes that authors require at least \onerep to post,
effectively blocking Sybil actions.
To vouch for new users, rule~\code{3.a} allows an existing user to like a
newbie's post to unblock it, but at the cost of \onerep.
This cost prevents that malicious members unblock new users indiscriminately,
which would be a breach for Sybils.
For the same reason, rule~\code{2} imposes a temporary cost of \onerep for
each new post.
%
Note that the pioneers rule~\code{1.a} solves the chicken-and-egg problem
imposed by rule~\code{4.a}: if new authors start with no \reps, but require
\reps to operate, it is necessary that some authors have initial \reps to boot
the chains.

@@

We performed experiments to evaluate the performance of \FC and its consensus
algorithm.
%
As detailed further, we measure the following evaluation parameters:
    (a) metadata overhead,
    (b) consensus runtime,
    (c) graph forks, and
    (d) blocked messages.
%
The main goal of this section is to stress the consensus algorithm to show that
permissionless public forums are viable, regardless of the inherent slower
performance in comparison to permissioned protocols.

@@

Evaluation item (b) accounts the consensus algorithm applied locally at each
chain.
We measured the time to sort posts in a local DAG both for the first time
(without any caches), and incrementally (from stable consensus caches).
%
For the chat chain, it takes \texttt{125s} and \texttt{50ms} for the initial
and incremental sorts respectively, while for the newsgroup chain, it takes
\texttt{100s} and \texttt{70ms}.
%
%Currently, we synchronize the DAGs takes hours because we the synchronization time
%
The incremental sort is limited to 7 days or 100 posts by design (as discussed
in Section~\ref{sec.design.hard}), regardless of the size of the chain, which
conveniently settles an upper bound on the input size of the consensus
algorithm.
%
Given that we use plain JSON files in the file system, we consider an
incremental consensus under \texttt{100ms} to be a practical upper bound.

@@

Decentralized topic-based publish-subscribe protocols, such as
    \emph{XMPP}~\cite{pubsub.xmpp},
    \emph{ActivityPub}~\cite{pubsub.activitypub}, and
    \emph{gossipsub}~\cite{pubsub.gossipsub},
decouples publishers from subscribers in the network.
%
A key limitation of \emph{pubsubs} is that the brokers that mediate
communication still have a special role in the network, such as authenticating
and validating posts.
%
Nevertheless, some pubsubs do not rely on server roles, and instead, use
P2P gossip dissemination~\cite{pubsub.tera,pubsub.rappel,pubsub.gossipsub}.
%,pubsub.stan,pubsub.vitis
Most of these protocols focus on techniques to achieve scalability and
performance, such as throughput, load balancing, and real-time relaying.
%
However, these techniques alone are not sufficient to operate permissionless
networks with malicious Sybils~\cite{pubsub.gossipsub2}.
Being generic protocols, pubsubs are typically unaware of the applications
built on top of them.
%
In contrast, as stated in Section~\ref{sec.freechains}, the pubsub of \FC is
conceptually at the application level and is integrated with the semantics of
chains, which already verifies posts at publishing time.
For instance, to flood the network with posts, malicious peers need to spend
reputation, which takes hours to recharge (rule~\code{2} in
Table~\ref{fig.rules}).
In addition, blocked posts (Figure~\ref{fig.state}) are not a concern either,
because they have limited reachability.
Another advantage of a tighter integration between the application and protocol
is that Merkle~DAGs simplify synchronization, provide persistence, and prevent
duplication of messages.
%
In summary, \FC and \emph{pubsub} middlewares operate at different network
layers, which suggests that \FC could benefit of the latter to manage the
interconnections between peers.

>>>>

In this work, we adapt the idea of scarce resources to reach consensus and
eradicate Sybils, but in the context of social content sharing.
We indicate 4 main contributions:
%
\begin{enumerate}
\item Recognizing published contents as scarce resources, since they require
human work.
Work is manifested as new posts, which if approved by others, reward authors
with reputation tokens, which are further used to evaluate other posts with
likes and dislikes.
With such \emph{proof-of-authoring} mechanism, token generation is expensive,
while verification is cheap and shared by multiple users.
%
\item Using the reputation system to determine consensus.
Due to decentralization and network partitions, posts in a timeline form a
causal graph with only partial order, which we promote to a total order based
on the reputation of authors.
The consensus order is fundamental to detect conflicting operations, such as
likes with insufficient reputation (akin to Bitcoin's double spending).
%
\item Allowing users to create diversified forums of interest (not a
singleton blockchain), each as an independent timeline with its own
subjective authoring etiquette.
%
\item Supporting content removal without compromising the integrity of the
decentralized blockchain.
Users have the power to revoke posts with dislikes, and peers are forced to
remove payloads, only forwarding associated blockchain metadata.
\end{enumerate}
%
We integrated the proposed consensus algorithm into \FC~\cite{fcs.sbseg20}, a
peer-to-peer (P2P) content dissemination protocol that provides strong eventual
consistency~\cite{p2p.sec}.
%p2p.crdts,
To show the feasibility of the consensus mechanism, we simulated a chat channel
and a newsgroup forum, both extracted from publicly available Internet archives.

@@

Without moderation, permissionless P2P forums are impractical
due to Sybils abusing the system.
For instance, it should take seconds to generate thousands of fake
identities and SPAM millions of messages.
%Hence, without moderation, there are no limits on the number and size of posts
%and no reasonable policy to distinguish quality.
For this reason, we propose a reputation system that works together with a
consensus algorithm to resist Sybils.
% and make peer-to-peer public forums practical.
We consider Sybils as groups of throw-away identities and machines with no
previous reputation in the forums.

@@

Note that even non-malicious partitions may experience conflicts when
merging.
For instance, dislike operations in one branch to past posts in the common
prefix could possibly block posts from users in the other branch.
%, depending on how the posts are ordered.
Therefore, we need to order branches with a global consensus, such that we can
validate operations consistently across the network.
%
Note also that, to reach consensus, we should preferably rely on the maximum
information that all branches agree on, which is exactly the posts in the
common prefix.

@@

As a counterpoint, maybe users $a$ and $b$ have abandoned the forum for months,
and thus \code{branch-1} is actually legitimate.
In this case, users $a$ and $b$ would succeed to take over the
forum, since they control the majority of \reps in the common
prefix.
Yet another possibility is that both branches are legitimate but became
disconnected for a long period.
In any case, it is unacceptable that an old remote branch affects a local
active forum.

For this reason, the consensus algorithm includes an extra constraint that
prevents long-lasting local branches to merge, leading to \emph{hard forks}.
A hard fork occurs when a local branch crosses a predetermined
threshold of \emph{7 days} or \emph{100 posts} of activity.
In this case, regardless of the remote branch reputation, the local branch
takes priority and is ordered first.
This situation is analogous to a hard fork in Bitcoin and the branches become
incompatible and will never synchronize again.
More than simple numeric disputes, a hard fork represents a social conflict in
which reconciling branches is no longer possible.

@@

In summary, the rules to merge a branch \code{j} from a remote machine into a
branch \code{i} from a local machine are as follows:
\begin{itemize}
    \item \code{i} is first if it crosses the activity threshold of
          \emph{7 days} or \emph{100 posts}, regardless of \code{j}.
    \item \code{i} or \code{j} is first, whichever has more reputation in the
          common prefix (Section~\ref{sec.design.basic}).
    \item otherwise, branches use the arbitrary criteria of lexicographical
          order of the post hashes immediately after the common prefix.
\end{itemize}

@@

In order to support content moderation and mitigate abuse, we integrated the
proposed reputation and consensus mechanism of Section~\ref{sec.design} with
\FC.
Table~\ref{fig.rules} details the reputation rules we conceived for
\FC, which is compatible with the general rules of Table~\ref{fig.general}.
%, and which are discussed as follows through an example.
Authors have to sign posts in order to be accounted by the reputation system
and operate in the chains.
We start by creating an identity whose public key is assigned as the pioneer in
a chain named \emph{forum}:

@@

%already in the community vouch for new users, or (ii) imposing explicit costs
%for new posts, such as proof of work.
%The most basic concern in public forums is to resist Sybil attacks.
%Fully P2P systems cannot rely on logins or {\footnotesize CAPTCHAs} due to the
%lack of a central authority.
%Viable alternatives include (i) building social trust graphs, in which users
%already in the community vouch for new users, or (ii) imposing explicit costs
%for new posts, such as proof of work.
%
%We propose a mix between trust graphs and economic costs.
%
In order to resist Sybils, we propose a mix between social trust
graphs and explicit costs for new posts:
Rule~\code{4.a} imposes to authors at least \onerep to post,
effectively blocking Sybil actions.
To vouch for new users, rule~\code{3.a} allows an existing user to like a
newbie's post to unblock it, but at the cost of \onerep.
This cost prevents malicious members to unblock new users indiscriminately.
%which would be a breach for Sybils.
For the same reason, rule~\code{2} imposes a temporary cost of \onerep for
each new post.
%
Note that the pioneers rule~\code{1.a} solves the chicken-and-egg problem
imposed by rule~\code{4.a}: if new authors start with no \reps, but require
\reps to operate, it is necessary that some authors have initial \reps to boot
the chains.

@@

We performed experiments to evaluate the performance of \FC and its consensus
algorithm.
%
As detailed next, we measure the following evaluation parameters:
    (a) metadata overhead,
    (b) consensus runtime,
    (c) graph forks, and
    (d) blocked messages.
%
Our goal is to stress the consensus algorithm to show that
permissionless public forums are viable, regardless of the inherent slower
performance in comparison to permissioned protocols.

@@

Evaluation item (b) accounts the consensus algorithm applied locally.
% at each chain.
We measured the time to sort posts in a local DAG both for the first time.
%(without any caches),
and incrementally from stable consensus caches.
%
For the chat chain, it takes \texttt{125s} and \texttt{50ms} for the initial
and incremental sorts,
% respectively,
while for the newsgroup chain, it takes
\texttt{100s} and \texttt{70ms}.
%
%Currently, we synchronize the DAGs takes hours because we the synchronization time
%
The incremental sort is limited to 7 days or 100 posts by design (as discussed
in Section~\ref{sec.design.hard}), regardless of the size of the chain, which
conveniently settles an upper bound on the input size of the consensus
algorithm.
%
Given that we use plain JSON files in the file system, we consider an
incremental consensus under \texttt{100ms} to be a practical upper bound.

@@

Decentralized topic-based publish-subscribe protocols, such as
    %\emph{XMPP}~\cite{pubsub.xmpp}, and
    \emph{ActivityPub}~\cite{pubsub.activitypub} and
    \emph{gossipsub}~\cite{pubsub.gossipsub},
decouples publishers from subscribers in the network.
%
A key limitation of \emph{pubsubs} is that the brokers that mediate
communication still have a special role in the network, such as authenticating
and validating posts.
%
Nevertheless, some pubsubs do not rely on server roles, and instead, use
P2P gossip dissemination~\cite{pubsub.tera,pubsub.rappel,pubsub.gossipsub}.
%,pubsub.stan,pubsub.vitis
Most of these protocols focus on techniques to achieve scalability and
performance, such as throughput, load balancing, and real-time relaying.
%
However, these techniques alone are not sufficient to operate permissionless
networks with malicious Sybils~\cite{pubsub.gossipsub2}.
Being generic protocols, pubsubs are typically unaware of the applications
built on top of them.
%
In contrast, as stated in Section~\ref{sec.freechains}, the pubsub of \FC is
conceptually at the application level and is integrated with the semantics of
chains, which already verifies posts at publishing time.
For instance, to flood the network with posts, malicious peers need to spend
reputation, which takes hours to recharge (rule~\code{2} in
Table~\ref{fig.rules}).
In addition, blocked posts (Figure~\ref{fig.state}) are not a concern either,
because they have limited reachability.
Another advantage of a tighter integration between the application and protocol
is that Merkle~DAGs simplify synchronization, provide persistence, and prevent
duplication of messages.
%
%In summary, \FC and \emph{pubsub} middlewares operate at different network
%layers, which suggests that \FC could benefit of the latter to manage the
%interconnections between peers.
